\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-25 paper.qmd
Add 3 cluster examples with suitable choice to discuss for PBMC3k

#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_5_min_dist_0.1.rds")
# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")

@@

## To find the benchmark value
benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
benchmark_pbmc <- 1.055271

@@

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.1) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/SO3thJ8PMi8>)"
#| label: fig-nldervisPBMCUMAP2
##| out-width: 100%
#| fig-pos: h

@@

#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_30_min_dist_0.01.rds")
# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")

@@

## To find the benchmark value
benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
benchmark_pbmc <- 1.057549

@@

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.01) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/hP8ie_oMXw0>)"
#| label: fig-nldervisPBMCUMAP3
##| out-width: 100%
#| fig-pos: h

@@

@@

plot_list2_harps <- tSNE_data_harps |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = cluster))+
    geom_point(alpha=0.5) +
    coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "cluster"),
         color = guide_legend(nrow = 3)) +
 annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

plot_list2_harps <- tSNE_data_harps |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = clusters))+
    geom_point(alpha=0.5) +
    coord_cartesian(xlim = c(-22, 18), ylim = c(-27, 32)) +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "cluster"),
         color = guide_legend(nrow = 3)) +
 annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  #coord_cartesian(xlim = c(-22, 18), ylim = c(-27, 32)) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from tSNE (perplexity = 23) applied for the HARPS GTO dataset. Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/_is9qStggFE>)"
#| label: fig-nldervisharpstSNE2
#| fig-pos: h

@@

trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  #coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from tSNE (perplexity = 91) applied for the HARPS GTO dataset. Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/aGvHFhWw_Xs>)"
#| label: fig-nldervisharpstSNE3
#| fig-pos: h

>>>>

#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_100_min_dist_0.3.rds")
# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")

@@

## To find the benchmark value
benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
benchmark_pbmc <- 1.049766

@@

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 100, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/1jNVVlzI4tM>)"
#| label: fig-nldervisPBMCUMAP2
##| out-width: 100%
#| fig-pos: h

@@

#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_15_min_dist_0.3.rds")
# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")

@@

## To find the benchmark value
benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
benchmark_pbmc <- 1.035666

@@

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/2GM_SuM866M>)"
#| label: fig-nldervisPBMCUMAP3
##| out-width: 100%
#| fig-pos: h

@@

@Anders2018

@@

plot_list2_harps <- tSNE_data_harps |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = cluster))+
    geom_point(alpha=0.5) +
    #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "cluster"),
         color = guide_legend(nrow = 3)) +
 annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

plot_list2_harps <- tSNE_data_harps |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = clusters))+
    geom_point(alpha=0.5) +
    coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "cluster"),
         color = guide_legend(nrow = 3)) +
 annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from tSNE (perplexity = 23) applied for the HARPS GTO dataset. Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/VO_4SJEoAUA>)"
#| label: fig-nldervisharpstSNE2
#| fig-pos: h

@@

trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from tSNE (perplexity = 91) applied for the HARPS GTO dataset. Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/vLvRfcogft8>)"
#| label: fig-nldervisharpstSNE3
#| fig-pos: h

### 2024-01-25 paper.tex
Add 3 cluster examples with suitable choice to discuss for PBMC3k

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-nldervisPBMCUMAP2}(a) 2D layout from UMAP
(n\_neighbors = 5, min\_dist = 0.1) applied for the PBMC3k dataset. Is
this a best representation of the original data?, (b) Model in the 2D
space with UMAP (\url{https://youtu.be/SO3thJ8PMi8})}

@@

\caption{\label{fig-nldervisPBMCUMAP3}(a) 2D layout from UMAP
(n\_neighbors = 30, min\_dist = 0.01) applied for the PBMC3k dataset. Is
this a best representation of the original data?, (b) Model in the 2D
space with UMAP (\url{https://youtu.be/hP8ie_oMXw0})}

@@

@@

\caption{\label{fig-nldervisharpstSNE2}(a) 2D layout from tSNE
(perplexity = 23) applied for the HARPS GTO dataset. Is this a best
representation of the original data?, (b) Model in the 2D space with
tSNE (\url{https://youtu.be/_is9qStggFE})}

@@

\caption{\label{fig-nldervisharpstSNE3}(a) 2D layout from tSNE
(perplexity = 91) applied for the HARPS GTO dataset. Is this a best
representation of the original data?, (b) Model in the 2D space with
tSNE (\url{https://youtu.be/aGvHFhWw_Xs})}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-nldervisPBMCUMAP2}(a) 2D layout from UMAP
(n\_neighbors = 100, min\_dist = 0.3) applied for the PBMC3k dataset. Is
this a best representation of the original data?, (b) Model in the 2D
space with UMAP (\url{https://youtu.be/1jNVVlzI4tM})}

@@

\caption{\label{fig-nldervisPBMCUMAP3}(a) 2D layout from UMAP
(n\_neighbors = 15, min\_dist = 0.3) applied for the PBMC3k dataset. Is
this a best representation of the original data?, (b) Model in the 2D
space with UMAP (\url{https://youtu.be/2GM_SuM866M})}

@@

\citet{Anders2018}

@@

\caption{\label{fig-nldervisharpstSNE2}(a) 2D layout from tSNE
(perplexity = 23) applied for the HARPS GTO dataset. Is this a best
representation of the original data?, (b) Model in the 2D space with
tSNE (\url{https://youtu.be/VO_4SJEoAUA})}

@@

\caption{\label{fig-nldervisharpstSNE3}(a) 2D layout from tSNE
(perplexity = 91) applied for the HARPS GTO dataset. Is this a best
representation of the original data?, (b) Model in the 2D space with
tSNE (\url{https://youtu.be/vLvRfcogft8})}
